\center \fontsize{18}{20} \selectfont Revealing the Landscape of Globally Color-Dual Multi-loop Integrands

### 2023-09-27 notes/2loopNLSM.tex
rest of edits

\title{\center  \fontsize{18.8}{20} \selectfont  Globally Color-Dual Multi-Loop Integrands}

\abstract{We report on progress in understanding how to construct color-dual multi-loop amplitudes.
First we identify a cubic theory, semi-abelian Yang-Mills, that unifies
  many of the color-dual theories studied in the literature, and
  provides a prescriptive approach for constructing $D$-dimensional color-dual
  numerators through one-loop directly from Feynman rules. By a simple weight counting argument, this approach does not further generalize to two-loops.  As a first step in
  understanding the two-loop challenge, we use a $D$-dimensional color-dual bootstrap to 
  successfully construct globally color-dual
  \emph{local} two-loop four-point nonlinear sigma model (NLSM) numerators.  The double-copy of the
these NLSM numerators with themselves, pure Yang-Mills, and
  $\mathcal{N}=4$ super-Yang-Mills correctly reproduce the known unitarity constructed integrands 
  of special Gallileons, Born-Infeld theory,
  and Dirac-Born-Infeld-Volkov-Akulov theory, respectively. 
  Applying our loop-level bootstrap, we exhaustively search the space of integrands consistent with point-like interactions and find that two-loop four-point pure Yang-Mills fails to satisfy global color-kinematics duality, completing a search previously initiated in the literature. 
  We pinpoint the failure to the
  bowtie unitarity cut, and discuss a path forward towards \emph{non-local}
  construction of color-dual integrands at generic loop order.

@@

\subsection{Semi-abelian Yang-Mills theory}\label{semiYM}
As we argued above, if we include factors of
$(\varepsilon\varepsilon)^{n\geq 1}$ at tree-level, then we need to
keep the Yang-Mills four-point contact for color-kinematics duality to
be restored on-shell. However, the contrapositive is also true -- if
we omit the four-point Yang-Mills vertex, then we only need terms that
contribute to the manifestly color-dual cubic sector of the theory.

We can achieve this construction by splitting the gauge group of the
full Yang-Mills action from $G$ to $G_L \times G_R$, where the full
non-abelian gauge symmetry is restored only when we identify
$G_L = G_R$. However, if we break the left gauge group down to
$G_L \rightarrow U(1)^{N^2}$ and take the right gauge group to
$G_R \rightarrow U(N)$, then we are left with a semi-abelian theory
that is gauge invariant in the preferred $U(1)$-directions. We call
this manifestly color-dual theory \textit{semi-abelian Yang-Mills},
\begin{eBox}
\begin{equation}
\label{eq:semiYM}
\mathcal{L}^{\text{semi-YM}} = -\frac{1}{2}\text{tr}\left[\bar{F}_{\mu\nu}F^{\mu\nu}\right]
\end{equation}
\end{eBox}
where
\begin{equation}
\bar{F}^a_{\mu\nu} = \partial_\mu \bar{A}_\nu^a - \partial_\nu \bar{A}_\mu^a
\end{equation}
\begin{equation}
F^a_{\mu\nu} = \partial_\mu A_\nu^a - \partial_\nu A_\mu^a + f^{abc} A_\mu^b A_\nu^c
\end{equation}
and $A_\mu$ is in Lorenz gauge.
Here we can think of the abelian vector as a background field that
sources on-shell $A_\mu$ currents. Indeed, as we describe in
\cref{sec:CKLagrangians}, semi-abelian Yang-Mills theory is at the
heart of $YZ$-theory \cite{Cheung:2016prv}, $J$-theory
\cite{Cheung:2020djz,Cheung:2021zvb}, and self-dual Yang-Mills
\cite{Monteiro2011pc}.  Specifically, semi-abelian YM is simply a
clever reinterpretation of the Lagrangian obtained by integrating an
auxiliary field into the $J$-theory equations of motion.  The Feynman
rule for the cubic vertex is
\begin{equation}
  \cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3)
  -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3)
\end{equation}
where the incoming arrow is the abelianized gauge field, and the outgoing arrows the on-shell non-abelian vectors. The propagator is simply
\be
\propFrule{$A^\nu$}{$\bar{A}^\mu$}{$k \rightarrow $}{fermion2} = \frac{i}{k^2} \eta^{\mu\nu}.
\ee
We can use the Ward identity of
\cref{eq:GIrelA} to check that the amplitudes are indeed gauge
invariant when the \emph{abelian} polarizations are taken to be longitudinal,
$\bar{\varepsilon} \rightarrow k$.
Furthermore, a simple calculation shows that the four-point correlation function of this theory satisfies the Jacobi identity of \cref{jacID} \emph{off-shell}. This ensures that color-kinematics duality holds to all multiplicity and loop order.
We also note that semi-YM does not have any ghosts from non-abelian gauge symmetry that could spoil color-kinematics at loop level. Due to the Feynman rules of this theory, the amplitudes are non-vanishing only at tree-level and one-loop. We demonstrate the implications of this property in \cref{2LoopObstruction}.

>>>>

\title{\center  \fontsize{18}{20} \selectfont  Revealing the Landscape of Globally Color-Dual Multi-loop Integrands}

\abstract{We report on progress in understanding how to construct color-dual multi-loop amplitudes.
First we identify a cubic theory, semi-abelian Yang-Mills, that unifies
  many of the color-dual theories studied in the literature, and
  provides a prescriptive approach for constructing $D$-dimensional color-dual
  numerators through one-loop directly from Feynman rules. By a simple weight counting argument, this approach does not further generalize to two-loops.  As a first step in
  understanding the two-loop challenge, we use a $D$-dimensional color-dual bootstrap to 
  successfully construct globally color-dual
  \emph{local} two-loop four-point nonlinear sigma model (NLSM) numerators.  The double-copy of the
these NLSM numerators with themselves, pure Yang-Mills, and
  $\mathcal{N}=4$ super-Yang-Mills correctly reproduce the known unitarity constructed integrands 
  of special Gallileons, Born-Infeld theory,
  and Dirac-Born-Infeld-Volkov-Akulov theory, respectively. 
  Applying our bootstrap to two-loop four-point pure Yang-Mills, we exhaustively search the space of local numerators and find that it fails to satisfy global color-kinematics duality, completing a search previously initiated in the literature. 
  We pinpoint the failure to the
  bowtie unitarity cut, and discuss a path forward towards \emph{non-local}
  construction of color-dual integrands at generic loop order.

@@

\subsection{Semi-abelian Yang-Mills theory}\label{semiYM}
As we argued above, if we include factors of
$(\varepsilon\varepsilon)^{n\geq 1}$ at tree-level, then we need to
keep the Yang-Mills four-point contact for color-kinematics duality to
be restored on-shell. However, the contrapositive is also true -- if
we omit the four-point Yang-Mills vertex, then we only need terms that
contribute to the manifestly color-dual cubic sector of the theory. We call
this manifestly color-dual theory \textit{semi-abelian Yang-Mills},
\begin{eBox}
\begin{equation}
\label{eq:semiYM}
\mathcal{L}^{\text{semi-YM}} = -\frac{1}{2}\text{tr}\left[\bar{F}_{\mu\nu}F^{\mu\nu}\right]
\end{equation}
\end{eBox}
where
\begin{equation}
\bar{F}^a_{\mu\nu} = \partial_\mu \bar{A}_\nu^a - \partial_\nu \bar{A}_\mu^a
\end{equation}
\begin{equation}
F^a_{\mu\nu} = \partial_\mu A_\nu^a - \partial_\nu A_\mu^a + f^{abc} A_\mu^b A_\nu^c
\end{equation}
and $A_\mu$ is in Lorenz gauge. We can construct this Lagrangian from \cref{eq:ym-lag} by keeping the right field strength covariant under
$U(N)$, and making the left field strength gauge covariant under
$U(1)^{N^2}$. Note that since $U(1)$ covariance is identical to $U(1)$ invariance, the amplitudes of this semi-abelian theory vanish under $\bar{\varepsilon}(k) \rightarrow k$, where $\bar{\varepsilon}$ is the polarization of an external $\bar{A}_\mu$ vector.

Here we can think of the abelian vector as a background field that
sources on-shell $A_\mu$ currents. Indeed, as we describe in
\cref{sec:CKLagrangians}, semi-abelian Yang-Mills theory is at the
heart of $YZ$-theory \cite{Cheung:2016prv}, $J$-theory
\cite{Cheung:2020djz,Cheung:2021zvb}, and self-dual Yang-Mills
\cite{Monteiro2011pc}.  Specifically, semi-abelian YM is simply a
clever reinterpretation of the Lagrangian obtained by integrating an
auxiliary field into the $J$-theory equations of motion.  The Feynman
rule for the cubic vertex is
\begin{equation}
  \cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3)
  -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3)
\end{equation}
where the incoming arrow is the abelianized gauge field, and the outgoing arrows the on-shell non-abelian vectors. The propagator is simply
\be
\propFrule{$A^\nu$}{$\bar{A}^\mu$}{$k \rightarrow $}{fermion2} = \frac{i}{k^2} \eta^{\mu\nu}.
\ee
We can use the Ward identity of
\cref{eq:GIrelA} to check that the amplitudes are indeed gauge
invariant when the \emph{abelian} polarizations are taken to be longitudinal,
$\bar{\varepsilon} \rightarrow k$.
Furthermore, a simple calculation shows that the four-point correlation function of this theory satisfies the Jacobi identity of \cref{jacID} \emph{off-shell}. This ensures that color-kinematics duality holds to all multiplicity and loop order.
We also note that semi-YM does not have any ghosts from non-abelian gauge symmetry that could spoil color-kinematics at loop level. Due to the Feynman rules of this theory, the amplitudes are non-vanishing only at tree-level and one-loop. We demonstrate the implications of this property in \cref{2LoopObstruction}.
